### 2014-11-23 fermithoughts.tex
adding fermi thoughts

\title{Fermi's Thoughts}
\date{}

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{amsthm}

\begin{document}
\maketitle

\section{Period 2}

\section{Period 3}

\section{Period 4}

\end{document}